---
title: "Football - EDA"
output: html_notebook
---

# Libraries

```{r}
library(dplyr)
library(lubridate)
library(DBI)
library(ggplot2)
library(zoo)
```


# Load data
```{r}
con = dbConnect(RSQLite::SQLite(), dbname="football_data_db.sqlite")
q <- "
  select *,
  case
      when length(date)>8 then substr(date,0,6)|| '/' || substr(date,9,2)
      else date
  end as repaired_date
  from football_data
  ;
"
df <- dbGetQuery(con, q)

dbDisconnect(con)
rm(con)
rm(q)
```


Date format
```{r}
df$Date <- as.Date(df$repaired_date, format = "%d/%m/%y")
```


Target Variable
```{r}
df <- df %>% mutate(target = as.numeric(FTHG>FTAG))
```



Filtering out proper observations
```{r}
dfm <- df %>% filter(
  !(Season %in% c("2004/2005", "2003/2004", "2002/2003")), 
  Country %in% c("Spain", "England", "Italy", "Germany", "Portugal", "France")
  )
```




# Counting Variables

## vars on match level

Points
  - h_points
  - a_points
Goals scored
  - FTHG
  - FTAG
Goals conceded
  - FTAG
  - FTHG
Goal balance
  - h_goal_balance
  - a_goal_balance
Corners
  - HC
  - AC
Corners balance
  - h_c_balance
  - a_c_balance
Shots
  - HS
  - AS
Shots balance
  - h_s_balance
  - a_s_balance
Shots on target
  - HST
  - AST
Shots on target balance
  - h_st_balance
  - a_st_balance
On Target Ratio
  - h_on_target_ratio
  - a_on_target_ratio
On Target Ratio Balance
  - h_on_target_ratio_balance
  - a_on_target_ratio_balance
Goal / Shots on target ratio
  - h_goals_st_ratio
  - a_goals_st_ratio
Goal / Shots on target ratio balance
  - h_goals_st_ratio_balance
  - a_goals_st_ratio_balance

  
Aggregate function - mean


h vs a 
diff
div


### Adding needed variables 

points in match
```{r}
dfm <- dfm %>% mutate(
  h_points = case_when(
    FTHG > FTAG ~ 3, 
    FTHG == FTAG ~ 1,
    TRUE ~ 0), 
  a_points = case_when(
    FTHG < FTAG ~ 3, 
    FTHG == FTAG ~ 1,
    TRUE ~ 0
  ),
  h_goal_balance = FTHG - FTAG, 
  a_goal_balance = FTAG - FTHG, 
  h_c_balance = HC - AC,
  a_c_balance = AC - HC,
  h_s_balance = HS - AS, 
  a_s_balance = AS - HS, 
  h_st_balance = HST - AST, 
  a_st_balance = AST - HST, 
  h_on_target_ratio = case_when(HS == 0 ~ 0, TRUE ~ HST/HS), 
  a_on_target_ratio = case_when(AS == 0 ~ 0, TRUE ~ AST/AS),
  h_on_target_ratio_balance = h_on_target_ratio - a_on_target_ratio,
  a_on_target_ratio_balance = a_on_target_ratio - h_on_target_ratio,
  h_goals_st_ratio = case_when(HST == 0 ~ 0, TRUE ~ FTHG/HST), 
  a_goals_st_ratio = case_when(AST == 0 ~ 0, TRUE ~ FTAG/AST),
  h_goals_st_ratio_balance = h_goals_st_ratio - a_goals_st_ratio,
  a_goals_st_ratio_balance = a_goals_st_ratio - h_goals_st_ratio
  
)
```


### Same but weighted by opponent score
```{r}
dfm$prob_H <- 1/dfm$B365H
dfm$prob_D <- 1/dfm$B365D
dfm$prob_A <- 1/dfm$B365A
  
dfm$margin <- dfm$prob_H + dfm$prob_D + dfm$prob_A - 1

dfm$real_prob_H <- dfm$prob_H / (1 + dfm$margin)
dfm$real_prob_D <- dfm$prob_D / (1 + dfm$margin)
dfm$real_prob_A <- dfm$prob_A / (1 + dfm$margin)
```


```{r}
vars_h <- c(
  "h_points",
  "FTHG",
  "FTAG",
  "h_goal_balance",
  "HC",
  "h_c_balance",
  "HS",
  "h_s_balance",
  "HST",
  "h_st_balance",
  "h_on_target_ratio",
  "h_on_target_ratio_balance",
  "h_goals_st_ratio",
  "h_goals_st_ratio_balance"
)

vars_a <- c(
  "a_points",
  "FTAG",
  "FTHG",
  "a_goal_balance",
  "AC",
  "a_c_balance",
  "AS",
  "a_s_balance",
  "AST",
  "a_st_balance",
  "a_on_target_ratio",
  "a_on_target_ratio_balance",
  "a_goals_st_ratio",
  "a_goals_st_ratio_balance"
)
```


home_team
```{r}
dfm <- dfm %>% mutate_at(.vars=vars(vars_h), .funs=funs(wgh_h = . * real_prob_A))
```

away team
```{r}
dfm <- dfm %>% mutate_at(.vars=vars(vars_a), .funs=funs(wgh_a = . * real_prob_H))
```




# Aggregation


## Home and Away number of matches played in a season

h_h_matches
```{r}
dfm <- dfm %>% arrange(Date) %>% group_by(Div, Season, HomeTeam) %>%
  mutate(h_h_matches=rollapply(HomeTeam, width = list(-1:-100), by=1, FUN = function(x){sum(!is.na(x))}, fill = NA, partial=T,align = "right"))
```


a_a_matches
```{r}
dfm <- dfm %>% arrange(Date) %>% group_by(Div, Season, AwayTeam) %>%
  mutate(a_a_matches=rollapply(AwayTeam, width = list(-1:-100), by=1, FUN = function(x){sum(!is.na(x))}, fill = NA, partial=T,align = "right"))
```

NA <- 0 
```{r}
dfm$h_h_matches[is.na(dfm$h_h_matches)] <- 0 
dfm$a_a_matches[is.na(dfm$a_a_matches)] <- 0
```


## Aggregation of match stats

```{r}
vars_h_all <- c(vars_h, paste(vars_h, "wgh_h", sep="_"))
vars_a_all <- c(vars_a, paste(vars_a, "wgh_a", sep="_"))
```


```{r}
dfm2 <- dfm %>% arrange(Date) %>% 
  
  # Home Team Aggregation
  group_by(Div, Season, HomeTeam) %>%
  mutate_at(vars(vars_h_all), funs(
    h_r3 = rollapply(data=., width = list(-1:-3), by=1, FUN = mean, fill = NA, na.rm = T, partial=T, align = "right")
    )) %>% ungroup() %>%
  
  # Away Team Aggrefation
  arrange(Date) %>%
  group_by(Div, Season, AwayTeam) %>%
  mutate_at(vars(vars_a_all), funs(
    a_r3 = rollapply(data=., width = list(-1:-3), by=1, FUN = mean, fill = NA, na.rm = T, partial=T, align = "right")
    )) %>% ungroup()
```



simple check

```{r}
pom <- dfm2 %>% filter(HomeTeam == "Arsenal") %>% arrange(Date) %>% select(Div, Date,h_h_matches, HomeTeam, AwayTeam, FTHG, FTAG, h_points, h_points_h_r3)
```


Save df to file

```{r}
dfm <- dfm2
rm(dfm2)
rm(pom)
```

```{r}
save(dfm, file="dfm.rdata")
```



# Ranking of variables
```{r}
library(randomForest)
```


```{r}
load(file="dfm.rdata")
```


```{r}

dfm2 <- dfm %>% filter(
  h_h_matches > 2,
  a_a_matches > 2,
  !is.na(real_prob_H), 
  !is.na(target)
)
```

```{r}
dfm2 <- dfm2[,c(156:211,100)]
```


```{r}
dfm3 <- dfm2[complete.cases(dfm2),]
```


```{r}
dfm3$target <- as.factor(dfm3$target)
```



```{r}
rfm <- randomForest(target ~ ., data=dfm3, importnace = TRUE)
```

Dataframe with Importance

```{r}
imp <- importance(rfm)
imp_df <- data.frame(variable = rownames(imp), MeanDecreaseGini = imp)
rownames(imp_df) <- NULL
colnames(imp_df) <- c("Variable", "MeanDecreaseGini")

imp_df <- imp_df[order(imp_df$MeanDecreaseGini, decreasing = T),]
```


Importance plot
```{r fig.height=12, fig.width=10}
library(caret)
varImpPlot(rfm,type=2)
```


Correlations

```{r fig.height=10, fig.width=10}
library(corrplot)
corrplot(cor(dfm3[,1:56]), order = "hclust", tl.cex=0.5)
```



