---
title: "Piłka nożna - regresja logistyczna"
output: html_notebook
---

Libraries
```{r}
library(dplyr)
library(ggplot2)
library(caret)
library(ROCR)
```


Load data
```{r}
setwd("~/Dysk_Google/datahacking/repos/bethacking")
load(file="dfm.rdata")
```


# Dealing with correlations 
```{r}
df_cor <- dfm[,c(18:73)]
```

```{r}
cor <- cor(df_cor)
```

```{r}
dissimilarity <- 1 - abs(cor)
distance <- as.dist(dissimilarity)
```

```{r fig.height=12, fig.width=10}
plot(hclust(distance), 
     main="Dissimilarity = 1 - abs(Correlation)", xlab="")
```

I choose cutoff equal to 0.8. 
It gives 11 groups of variables. I will try to choose one representative for each group. 

```{r}
hc <- hclust(distance)
```

```{r}
clusterCut <- cutree(hc, 11)
```

```{r}
clusters_df <- data.frame(variable = names(clusterCut), cluster = clusterCut, stringsAsFactors = F)
row.names(clusters_df) <- NULL
```

now we need ranking of variables. I will use random forest to generate it. 

```{r}
library(randomForest)
set.seed(2018)
rfm <- randomForest(as.factor(target) ~ ., dfm[,c(18:74)], importance=TRUE)
```

```{r}
var_rank <- data.frame(variable = row.names(importance(rfm)), MeanDecreaseGini=importance(rfm)[,4], stringsAsFactors = FALSE)
row.names(var_rank) <- NULL
```

join clusters with ranking
```{r}
cl_rnk <- merge(clusters_df, var_rank, by="variable")
cl_rnk <- cl_rnk %>% arrange(cluster)
```

Choose max decrease gini for every cluster
```{r}
cl_rnk <- cl_rnk %>% group_by(cluster) %>% mutate(max_cl_imp = max(MeanDecreaseGini)) %>% ungroup() %>%
  filter(max_cl_imp == MeanDecreaseGini)
```


# Train / Valid Split
```{r}
set.seed(2018)
train_ind <- sample(seq_len(nrow(dfm)), size = nrow(dfm)/2)

tr <- dfm[train_ind,]
va <- dfm[-train_ind,]

```

# Logistic Regression Model 
```{r}
tr <- tr[,c("target", cl_rnk$variable)]
va <- va[,c("target", cl_rnk$variable)]
lrm <- glm(as.factor(target) ~ ., data = tr, family = "binomial")
summary(lrm)
```

```{r}
tr$AC_wgh_a_a_r3 <- NULL
tr$h_on_target_ratio_h_r3 <- NULL
tr$lr_prob <- NULL
tr$lr_forecast <- NULL
lrm <- glm(as.factor(target) ~ ., data = tr, family = "binomial")
summary(lrm)

```

Train
```{r}
tr$lr_prob <- predict(lrm, newdata=tr, type="response")
tr$lr_forecast <- as.numeric(tr$lr_prob > .5)

va$lr_prob <- predict(lrm, newdata=va, type="response")
va$lr_forecast <- as.numeric(va$lr_prob > .5)
```


Confusion Matrix - Train Set
```{r}
confusionMatrix(tr$lr_forecast, tr$target, positive = '1')
```

Confusion Matrix - Validation Set
```{r}
confusionMatrix(va$lr_forecast, va$target, positive = '1')
```


Training Set ROC AUC
```{r}
library(ROCR)
pred <- prediction(tr$lr_prob, tr$target)
perf <- performance(pred, measure = "auc")
perf@y.values
```


Validation Set ROC AUC
```{r}
library(ROCR)
pred <- prediction(va$lr_prob, va$target)
perf <- performance(pred, measure = "auc")
perf@y.values
```



Correlation Matrix
```{r fig.height=5, fig.width=5}
tr_cor <- tr[,c(2:10)]
library(corrplot)
corrplot(cor(tr_cor), order = "hclust", tl.cex=0.5)
```

